---
title: ' Projets tutorés 2 | Devoir 3'
author: "YAHIA Mohammed Rafik"
date: "2023-05-17"
output:
  prettydoc::html_pretty:
    theme: cayman
    number_sections: true
    toc: true
    toc_depth: 3
    df_print: paged
    css: rafik.css
    highlight : github
  pdf_document: default
---

<style>
.table-hover > tbody > tr:hover { 
  background-color: #e1f1fd;
}
</style>

```{css, echo=FALSE}
pre {
  max-height: 200px;
  overflow-y: auto;
}

pre[class] {
  max-height: 80px;
}
```

```{r setup, include=FALSE, echo=FALSE}
require(tidyverse)
require(caret)
require(missForest)
require(explore)
require(plyr)
require(janitor)
require(lfactors)
require(magrittr)
require(caretEnsemble)
require(doParallel)
require(skimr)
require(kableExtra)
require(tibble)
require(stringr)
require(prettydoc)
require(rmdformats)

#########################################################
### DM Table with variable that indicates the ARM. LB Table that indicates drug results and day.
dm <- read.csv("ascii-data-files nida-ctn-0001-20230318/dm.csv", header = T, dec = ".")
dm2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/dm.csv", header = T, dec = ".")
lb <- read.csv("ascii-data-files nida-ctn-0001-20230318/lb.csv", header = T, dec = ".")
lb2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/lb.csv", header = T, dec = ".")

#Criteria of success D13 or D14 with negative opoids results.
# Importing lb database containg opioid test results, and dm with arm.
dm3 <- dm %>% filter(ARM %in% c("CLONIDINE","BUPRENORPHINE/NALOXONE")) %>% select(USUBJID, ARM)
dm4 <- dm2 %>% filter(ARM %in% c("CLONIDINE","BUPRENORPHINE/NALOXONE")) %>% select (USUBJID, ARM)
lb3 <- lb %>% filter(EPOCH == "ACTIVE", VISITNUM %in% c(1:14)) #Missing 5 patients from lb3 compared to dm3 when applying this filter. as they are still in screening epoch but have already been assigned in dm database "01_018076" "01_027452" "01_058918" "01_063935" "01_067193".
lb4 <- lb2 %>% filter(EPOCH == "ACTIVE", VISITNUM %in% c(1:14)) #Missing 27 patients from lb4 compared to dm4 when applying this filter.

# Adding the ARM column.
list_df10 = list(lb3, dm3)
tabURINE <- list_df10 %>% 
  reduce(full_join, by='USUBJID') %>% 
  select("USUBJID", "LBTEST", "LBORRES", "LBSTAT", "LBREASND", "LBSPEC", "LBMETHOD", "VISIT", "VISITNUM", "LBDY", "LBDTC", "ARM") #113 patients

list_df11 = list(lb4, dm4)
tabURINE2 <- list_df11 %>% 
  reduce(full_join, by='USUBJID') %>% 
  select("USUBJID", "LBTEST", "LBORRES", "LBSTAT", "LBREASND", "LBSPEC", "LBMETHOD", "VISIT", "VISITNUM", "LBDY", "LBDTC", "ARM") #231 patients

# Merging both tables
TableURINE <- rbind(tabURINE, tabURINE2)

# Success criteria met for these patients with NEGATIVE test results, others are a Failure.
TableSuc <- TableURINE %>% 
  filter(VISITNUM %in% c(13,14) ) %>% 
  filter(LBTEST == "OPIATE") %>% 
  filter(LBORRES == "NEGATIVE") %>% 
  mutate(PROTOCOLE = if_else(str_detect(USUBJID, "^01"), "In-patient protocol", "Out-patient protocol")) %>%
  mutate(PROTOCOLE = as.factor(PROTOCOLE)) %>% 
  filter(ARM %in% c("CLONIDINE","BUPRENORPHINE/NALOXONE")) %>% 
  # Issue with patient 02_036130. Filtered again because after join there was one patient that had SCREEN FAILURE on DM4 but got to the final table through LB4 because he had active epoch tests, which doesn't make sense.
  arrange(VISITNUM) %>% 
  select(USUBJID, LBORRES)

# TableAgg from Devoir 1, for more information please refer to Devoir 1.
dm <- read.csv("ascii-data-files nida-ctn-0001-20230318/dm.csv", header = T, dec = ".")
tabDM <- dm %>% select(USUBJID, ARM, AGE, SEX)
vs <- read.csv("ascii-data-files nida-ctn-0001-20230318/vs.csv", header = T, dec = ".")
tabTEMP <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="TEMP") %>% select(USUBJID, VSORRES)
tabTEMP <- distinct(tabTEMP)
colnames(tabTEMP) <- c("USUBJID", "TEMP")
tabRESP <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="RESP") %>% select(USUBJID, VSORRES)
tabRESP <- distinct(tabRESP)
colnames(tabRESP) <- c("USUBJID", "RESP")
tabPULSEsitting <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEsitting <- distinct(tabPULSEsitting)
colnames(tabPULSEsitting) <- c("USUBJID", "PULSEsitting")
tabPULSEstanding <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEstanding <- distinct(tabPULSEstanding)
colnames(tabPULSEstanding) <- c("USUBJID", "PULSEstanding")
tabSBPsitting <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPsitting <- distinct(tabSBPsitting)
colnames(tabSBPsitting) <- c("USUBJID", "SBPsitting")
tabSBPstanding <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPstanding <- distinct(tabSBPstanding)
colnames(tabSBPstanding) <- c("USUBJID", "SBPstanding")
tabDBPsitting <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPsitting <- distinct(tabDBPsitting)
colnames(tabDBPsitting) <- c("USUBJID", "DBPsitting")
tabDBPstanding <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPstanding <- distinct(tabDBPstanding)
colnames(tabDBPstanding) <- c("USUBJID", "DBPstanding")
tabHEIGHT <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="HEIGHT") %>% select(USUBJID, VSORRES)
tabHEIGHT <- distinct(tabHEIGHT)
colnames(tabHEIGHT) <- c("USUBJID", "HEIGHT") 
tabWEIGHT <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="WEIGHT") %>% select(USUBJID, VSORRES)
tabWEIGHT <- distinct(tabWEIGHT)
colnames(tabWEIGHT) <- c("USUBJID", "WEIGHT")
calculate_bmi <- function(dataframe, height_col, weight_col, bmi_col) {
  height_meters <- dataframe[[height_col]] * 0.0254
  weight_kg <- dataframe[[weight_col]] * 0.453592 
  bmi <- weight_kg / (height_meters^2)
  dataframe[[bmi_col]] <- bmi 
  return(dataframe)
}
tabBMI <- left_join(tabHEIGHT, tabWEIGHT, by="USUBJID")
tabBMI <- calculate_bmi(tabBMI, "HEIGHT", "WEIGHT", "BMI") %>% select("USUBJID", "BMI")
list_df1 = list(tabBMI, tabDBPsitting, tabSBPsitting, tabDBPstanding, tabSBPstanding, tabPULSEsitting, tabPULSEstanding, tabRESP, tabTEMP)
tabVS <- list_df1 %>% reduce(full_join, by='USUBJID')
sc <- read.csv("ascii-data-files nida-ctn-0001-20230318/sc.csv", header = T, dec = ".")
tabMARITAL <- sc %>% filter(SCTESTCD=="MARITAL") %>% select(USUBJID, SCORRES)
tabMARITAL <- distinct(tabMARITAL)
colnames(tabMARITAL) <- c("USUBJID", "MARITAL")
tabEDUC <- sc %>% filter(SCTESTCD=="EDUCYRS") %>% select(USUBJID, SCORRES)
tabEDUC <- distinct(tabEDUC)
colnames(tabEDUC) <- c("USUBJID", "EDUCYRS")
tabEMPLOY30 <- sc %>% filter(SCTESTCD=="EMPLOY30") %>% select(USUBJID, SCORRES)
tabEMPLOY30 <- distinct(tabEMPLOY30)
colnames(tabEMPLOY30) <- c("USUBJID", "EMPLOY30")
tabEMPLOY3Y <- sc %>% filter(SCTESTCD=="EMPLOY3Y") %>% select(USUBJID, SCORRES)
tabEMPLOY3Y <- distinct(tabEMPLOY3Y)
colnames(tabEMPLOY3Y) <- c("USUBJID", "EMPLOY3Y")
tabEMPLOY <- left_join(tabEMPLOY30, tabEMPLOY3Y, by="USUBJID")
list_df2 = list(tabEDUC, tabEMPLOY, tabMARITAL)
tabSC <- list_df2 %>% reduce(full_join, by='USUBJID')
lb <- read.csv("ascii-data-files nida-ctn-0001-20230318/lb.csv", header = T, dec = ".")
tabMETHADONE <- lb %>% filter(EPOCH=="SCREENING", LBTESTCD=="METHADON", LBMETHOD=="CENTRAL LAB") %>% select(USUBJID, LBORRES)
tabMETHADONE <- distinct(tabMETHADONE)
colnames(tabMETHADONE) <- c("USUBJID", "METHADONE")
tabMORPHINE <- lb %>% filter(EPOCH=="SCREENING", LBTESTCD=="MORPHINE") %>% select(USUBJID, LBORRES)
tabMORPHINE <- distinct(tabMORPHINE)
colnames(tabMORPHINE) <- c("USUBJID", "MORPHINE")
list_df3 = list(tabMETHADONE, tabMORPHINE)
tabLB <- list_df3 %>% reduce(full_join, by='USUBJID')
qs <- read.csv("ascii-data-files nida-ctn-0001-20230318/qs.csv", header = T, dec = ".")
tabQS <- qs %>% filter(EPOCH=="SCREENING", QSTESTCD=="VAS001") %>% select(USUBJID, QSORRES)
tabQS <- distinct(tabQS)
colnames(tabQS) <- c("USUBJID", "VAS001")
list_df4 = list(tabDM, tabSC, tabVS, tabQS, tabLB)
tabALL_01 <- list_df4 %>% reduce(full_join, by='USUBJID')
dm2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/dm.csv", header = T, dec = ".")
tabDM2 <- dm2 %>% select(USUBJID, ARM, AGE, SEX)
vs2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/vs.csv", header = T, dec = ".")
tabTEMP2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="TEMP") %>% select(USUBJID, VSORRES)
tabTEMP2 <- distinct(tabTEMP2)
colnames(tabTEMP2) <- c("USUBJID", "TEMP")
tabTEMP2$TEMP[tabTEMP2$TEMP == 999.9] <- 99.9 
tabRESP2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="RESP") %>% select(USUBJID, VSORRES)
tabRESP2 <- distinct(tabRESP2)
colnames(tabRESP2) <- c("USUBJID", "RESP")
tabPULSEsitting2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEsitting2 <- distinct(tabPULSEsitting2)
colnames(tabPULSEsitting2) <- c("USUBJID", "PULSEsitting")
tabPULSEstanding2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEstanding2 <- distinct(tabPULSEstanding2)
colnames(tabPULSEstanding2) <- c("USUBJID", "PULSEstanding")
tabSBPsitting2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPsitting2 <- distinct(tabSBPsitting2)
colnames(tabSBPsitting2) <- c("USUBJID", "SBPsitting")
tabSBPstanding2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPstanding2 <- distinct(tabSBPstanding2)
colnames(tabSBPstanding2) <- c("USUBJID", "SBPstanding")
tabDBPsitting2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPsitting2 <- distinct(tabDBPsitting2)
colnames(tabDBPsitting2) <- c("USUBJID", "DBPsitting")
tabDBPstanding2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPstanding2 <- distinct(tabDBPstanding2)
colnames(tabDBPstanding2) <- c("USUBJID", "DBPstanding")
tabHEIGHT <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="HEIGHT") %>% select(USUBJID, VSORRES)
tabHEIGHT <- distinct(tabHEIGHT)
colnames(tabHEIGHT) <- c("USUBJID", "HEIGHT")
tabWEIGHT <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="WEIGHT") %>% select(USUBJID, VSORRES)
tabWEIGHT <- distinct(tabWEIGHT)
colnames(tabWEIGHT) <- c("USUBJID", "WEIGHT")
tabHEIGHT2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="HEIGHT") %>% select(USUBJID, VSORRES)
tabHEIGHT2 <- distinct(tabHEIGHT2)
colnames(tabHEIGHT2) <- c("USUBJID", "HEIGHT")
tabHEIGHT2$HEIGHT[tabHEIGHT2$HEIGHT == 715] <- 71
tabWEIGHT2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="WEIGHT") %>% select(USUBJID, VSORRES)
tabWEIGHT2 <- distinct(tabWEIGHT2)
colnames(tabWEIGHT2) <- c("USUBJID", "WEIGHT")
calculate_bmi <- function(dataframe, height_col, weight_col, bmi_col) {
  height_meters <- dataframe[[height_col]] * 0.0254
  weight_kg <- dataframe[[weight_col]] * 0.453592 
  bmi <- weight_kg / (height_meters^2)  
  dataframe[[bmi_col]] <- bmi
  return(dataframe)
}
tabBMI2 <- left_join(tabHEIGHT2, tabWEIGHT2, by="USUBJID")
tabBMI2 <- calculate_bmi(tabBMI2, "HEIGHT", "WEIGHT", "BMI") %>% select("USUBJID", "BMI")
list_df5 = list(tabBMI2, tabDBPstanding2, tabSBPstanding2, tabDBPsitting2, tabSBPsitting2, tabPULSEstanding2, tabPULSEsitting2, tabRESP2, tabTEMP2)
tabVS2 <- list_df5 %>% reduce(full_join, by='USUBJID')
sc2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/sc.csv", header = T, dec = ".")
tabMARITAL2 <- sc2 %>% filter(SCTESTCD=="MARITAL") %>% select(USUBJID, SCORRES)
tabMARITAL2 <- distinct(tabMARITAL2)
colnames(tabMARITAL2) <- c("USUBJID", "MARITAL")
tabEDUC2 <- sc2 %>% filter(SCTESTCD=="EDUCYRS") %>% select(USUBJID, SCORRES)
tabEDUC2 <- distinct(tabEDUC2)
colnames(tabEDUC2) <- c("USUBJID", "EDUCYRS")
tabEMPLOY302 <- sc2 %>% filter(SCTESTCD=="EMPLOY30") %>% select(USUBJID, SCORRES)
tabEMPLOY302 <- distinct(tabEMPLOY302)
colnames(tabEMPLOY302) <- c("USUBJID", "EMPLOY30")
tabEMPLOY3Y2 <- sc2 %>% filter(SCTESTCD=="EMPLOY3Y") %>% select(USUBJID, SCORRES)
tabEMPLOY3Y2 <- distinct(tabEMPLOY3Y2)
colnames(tabEMPLOY3Y2) <- c("USUBJID", "EMPLOY3Y")
tabEMPLOY2 <- left_join(tabEMPLOY302, tabEMPLOY3Y2, by="USUBJID")
list_df6 = list(tabEDUC2, tabEMPLOY2, tabMARITAL2)
tabSC2 <- list_df6 %>% reduce(full_join, by='USUBJID')
lb2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/lb.csv", header = T, dec = ".")
tabMETHADONE2 <- lb2 %>% filter(EPOCH=="SCREENING", LBTESTCD=="METHADON", LBMETHOD=="CENTRAL LAB") %>% select(USUBJID, LBORRES)
tabMETHADONE2 <- distinct(tabMETHADONE2)
colnames(tabMETHADONE2) <- c("USUBJID", "METHADONE")
tabMORPHINE2 <- lb2 %>% filter(EPOCH=="SCREENING", LBTESTCD=="MORPHINE") %>% select(USUBJID, LBORRES)
tabMORPHINE2 <- distinct(tabMORPHINE2)
colnames(tabMORPHINE2) <- c("USUBJID", "MORPHINE") 
list_df7 = list(tabMETHADONE2, tabMORPHINE2)
tabLB2 <- list_df7 %>% reduce(full_join, by='USUBJID')
qs2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/qs.csv", header = T, dec = ".")
tabQS2 <- qs2 %>% filter(EPOCH=="SCREENING", QSTESTCD=="VAS001") %>% select(USUBJID, QSORRES)
tabQS2 <- distinct(tabQS2)
colnames(tabQS2) <- c("USUBJID", "VAS001")
list_df8 = list(tabDM2, tabSC2, tabVS2, tabQS2, tabLB2)
tabALL_02 <- list_df8 %>% reduce(full_join, by='USUBJID')
TableALL <- rbind(tabALL_02, tabALL_01)
TableALL$VAS001 <- as.numeric(TableALL$VAS001)

TableAgg <- TableALL %>%
  group_by(USUBJID, ARM, SEX, EDUCYRS, EMPLOY30, EMPLOY3Y, MARITAL, METHADONE, MORPHINE) %>%
  dplyr::summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>% 
  mutate(dplyr::across(everything(), ~replace_na(.x, NA)))

TableAgg <- distinct(TableAgg) %>% 
  filter(!(ARM == "")) %>% 
  filter(!(ARM == "SCREEN FAILURE"))
TableAgg$ARM <- as.factor(TableAgg$ARM)
TableAgg$SEX <- as.factor(TableAgg$SEX)
TableAgg$EDUCYRS <- as.numeric(TableAgg$EDUCYRS)
TableAgg$EMPLOY30 <- as.factor(TableAgg$EMPLOY30)
TableAgg$EMPLOY3Y <- as.factor(TableAgg$EMPLOY3Y)
TableAgg$MARITAL <- as.factor(TableAgg$MARITAL)
TableAgg$METHADONE <- as.factor(TableAgg$METHADONE)
TableAgg$MORPHINE <- as.factor(TableAgg$MORPHINE)
TableAgg <- as.data.frame(TableAgg) 


# Merging this table with the initial table we have from Devoir 1 
# Now we have the table with Success 1 or Failure 0 in LBORRES column.
list_df20 = list(TableAgg, TableSuc)
Dataset_01 <- list_df20 %>% 
  reduce(full_join, by='USUBJID') %>% 
  mutate(PROTOCOLE = if_else(str_detect(USUBJID, "^01"), "In-patient protocol", "Out-patient protocol")) %>%
  mutate(PROTOCOLE = as.factor(PROTOCOLE)) %>% 
  mutate(LBORRES = if_else(is.na(LBORRES), "0", LBORRES)) %>% 
  mutate(LBORRES = if_else(str_detect(LBORRES, "NEGATIVE"), "1", "0")) %>% 
  mutate(LBORRES = as.factor(LBORRES)) %>% 
  mutate(EDUCYRS = as.factor(EDUCYRS))

# EVA
# Trends to check : Protocole, arm, sex & age effect on Treatement results LBORRES
# explore(DS01_norm)

# Preparing the dataset to apply ML algorithms.
# filling NA, One hot encoding, Normalizing
skim(Dataset_01)
Dataset_01_treated <- Dataset_01 %>% 
  select(LBORRES, USUBJID, PROTOCOLE, ARM, AGE, SEX, BMI, METHADONE, MORPHINE, VAS001, EDUCYRS, EMPLOY30, EMPLOY3Y, MARITAL, DBPsitting, SBPsitting, PULSEsitting, RESP, TEMP) %>% 
  mutate(PROTOCOLE = if_else(str_detect(PROTOCOLE, "^In"), "0", "1")) %>% 
  mutate(PROTOCOLE = as.factor(PROTOCOLE)) %>% 
  mutate(ARM = if_else(str_detect(ARM, "^BUP"), "0", "1")) %>% 
  mutate(ARM = as.factor(ARM)) %>% 
  mutate(SEX = if_else(str_detect(SEX, "F"), "0", "1")) %>% 
  mutate(SEX = as.factor(SEX)) %>% 
  mutate(METHADONE = if_else(str_detect(METHADONE, "^NEG"), "0", "1")) %>% 
  mutate(METHADONE = as.factor(METHADONE)) %>% 
  mutate(MORPHINE = if_else(str_detect(MORPHINE, "^NEG"), "0", "1")) %>% 
  mutate(MORPHINE = as.factor(MORPHINE))

# Removing NA # https://www.kaggle.com/code/lmorgan95/missforest-the-best-imputation-algorithm/report
set.seed(70)
Dataset_IMP <- subset(Dataset_01_treated, select = -c(LBORRES, USUBJID))
imp_missForest <- missForest(Dataset_IMP, maxiter = 10, verbose = T)
Dataset_IMP <- imp_missForest$ximp
# USUBJID Not needed as we already extracted arm information in a seperate column
LBORRES <- Dataset_01_treated[, 1]
Dataset_01_treated <- cbind(LBORRES, Dataset_IMP)

# Normalizing values 
preproc <- preProcess(Dataset_01_treated, method = "range")
DS01_norm <- predict(preproc, Dataset_01_treated)

#DS01_norm$LBORRES <- revalue(DS01_norm$LBORRES, c("0"="Zero", "1"="One"))
#DS01_norm$LBORRES <- relevel(DS01_norm$LBORRES, ref = "One")

# Dummy variables
dmy <- dummyVars(" ~ .", data = DS01_norm, fullRank=T)
DS01_norm_dummy <- data.frame(predict(dmy, newdata = DS01_norm)) #FullRank to choose one dummy for columns that has 2 values, ex GENDER.
DS01_norm_dummy <- DS01_norm_dummy %>% clean_names()
colnames(DS01_norm_dummy)[colnames(DS01_norm_dummy) == "lborres_1"] <- "Treatment_Results"
DS01_norm_dummy$Treatment_Results <- lfactor(DS01_norm_dummy$Treatment_Results, levels = 0:1, labels = c("Failure", "Success"))
DS01_norm_dummy$Treatment_Results <- as.factor(DS01_norm_dummy$Treatment_Results)

set.seed(70)
index <- createDataPartition(DS01_norm_dummy$Treatment_Results, p = 0.80, list = FALSE)
train <- DS01_norm_dummy[index,]
test <- DS01_norm_dummy[-index,]
testref <- DS01_norm_dummy[-index,]


TableSKIMR <- DS01_norm_dummy %>% 
  skim() %>% 
  dplyr::select(-skim_type)

















# If our prediction are all "Failure" we get an Accuracy of 69% (Sens 0 no positive class)
# This should be the baseline we improve upon in both Accuracy and Sensitivity.
test$Treatment_Results = "Failure"
test$Treatment_Results <- lfactor(test$Treatment_Results, levels = 0:1, labels = c("Failure", "Success"))
confusionMatrix(data = test$Treatment_Results, reference = testref$Treatment_Results, positive = "Success")

## Feature selection using RFE (Recursive Feature Elimination)
# Selected variable : ARM, PROTOCOLE
set.seed(70)
Control <- rfeControl(functions=rfFuncs, method="cv", number=10) 
Results <- rfe(DS01_norm[,2:18], DS01_norm[,1], sizes=c(1:16), metric = "Kappa", rfeControl=Control)
print(Results)
predictors(Results)
plot(Results)

## Feature selection using SBF (Selection By Filtering)
# Selected variable : ARM (100%), MORPHINE (100%), PROTOCOLE (100%), SBPsitting (100%), PULSEsitting (60%)
DS01_norm$LBORRES <- as.numeric(DS01_norm$LBORRES)
DS01_norm$PROTOCOLE <- as.numeric(DS01_norm$PROTOCOLE)
DS01_norm$ARM <- as.numeric(DS01_norm$ARM)
DS01_norm$SEX <- as.numeric(DS01_norm$SEX)
DS01_norm$METHADONE <- as.numeric(DS01_norm$METHADONE)
DS01_norm$MORPHINE <- as.numeric(DS01_norm$MORPHINE)

numeric_cols <- DS01_norm %>% select_if(is.numeric)

numeric_cols$LBORRES <- numeric_cols$LBORRES - 1
numeric_cols$PROTOCOLE <- numeric_cols$PROTOCOLE - 1 
numeric_cols$ARM <- numeric_cols$ARM - 1
numeric_cols$SEX <- numeric_cols$SEX - 1
numeric_cols$METHADONE <- numeric_cols$METHADONE - 1
numeric_cols$MORPHINE <- numeric_cols$MORPHINE -1

numeric_cols$LBORRES <- as.factor(numeric_cols$LBORRES)


RFwithGAM <- sbf(numeric_cols[,2:14], numeric_cols[,1],
                 sbfControl = sbfControl(functions = rfSBF,
                                         verbose = FALSE,
                                         method = "cv"))
RFwithGAM


## Algorithm selection
# https://datascience.stackexchange.com/questions/10745/which-of-the-180-algorithms-in-rs-caret-package-are-feasible
all_model <- modelLookup()
classification_model <- all_model %>% 
  filter(forClass == TRUE, !duplicated(model))

# Packages that will be used for training these models: 
all_packages <- sapply(classification_model$model, 
                       function(x) {
                         x %>% getModelInfo() %>% .[[1]] %>% .[["library"]]}) %>% unlist() 

all_packages <- all_packages[!duplicated(all_packages)]

your_packages <- installed.packages() %>% 
  as.data.frame() %>% 
  pull(Package) %>% 
  as.character()

cannot_installed <- c("adaptDA", "CHAID", "sparsediscrim", "elmNN", "gpls", 
                      "logicFS", "FCNN4R", "mxnet", "vbmp")

ml_not_used <- c("amdai", "chaid", "dda", "elm", "gpls", 
                 "logicBag", "mlpSGD1", "mxnet", "vbmpRadial")

registerDoParallel(cores = detectCores() - 1)
additional_package <- all_packages[!all_packages %in% your_packages]
additional_package <- all_packages[!all_packages %in% cannot_installed]
additional_package %>% length()
should_use_ml <- classification_model %>% 
  filter(!model %in% ml_not_used)
all_model <- should_use_ml$model
my_models <- all_model[c(1,51, 52, 130, 161, 182)]

# Training multiple models
number <- 5
repeats <- 4
control <- trainControl(method = "cv", 
                        classProbs = TRUE, 
                        savePredictions = "final",
                        index = createResample(train$Treatment_Results, repeats*number), 
                        summaryFunction = multiClassSummary, 
                        allowParallel = TRUE)
set.seed(70)

model_list1 <- caretList(Treatment_Results ~ protocole_1+arm_1+morphine_1+sb_psitting+puls_esitting, 
                         data = train,
                         trControl = control,
                         metric = "Accuracy", 
                         methodList = my_models)

# Plotting with accuracy the different models: 
list_of_results <- lapply(my_models, function(x) {model_list1[[x]]$resample})
df_results <- do.call("bind_rows", list_of_results)
df_results %<>% mutate(Model = lapply(my_models, function(x) {rep(x, number*repeats)}) %>% unlist())
df_results %>% 
  select(Accuracy, Model) %>% 
  ggplot(aes(Model, Accuracy, fill = Model, color = Model)) + 
  geom_boxplot(show.legend = FALSE, alpha = 0.3) + 
  theme_minimal() + 
  coord_flip()



## Random forest
# All features resulting in not so optimal results. (ACC, SENS, SPEC - 76%, 38%, 93%)
# With features selection, we get optimal results. (ACC, SENS, SPEC - 80%, 47%, 95%)
RFmodel <- predict(model_list1$rf, newdata = test, type="raw")
model_list1$rf #73% accuracy, difference between train accuracy and test acc is the overfitting
confusionMatrix(data = RFmodel, reference = testref$Treatment_Results, positive = "Success")


## Logistic regression
# All features. (ACC, SENS, SPEC - 72%, 28%, 91%)
# Feature selec. (ACC, SENS, SPEC - 78%, 42%, 93%)
set.seed(70)
LRmodel <- predict(model_list1$glm, newdata = test, type="raw")
model_list1$glm #77% training accuracy
confusionMatrix(data = LRmodel, reference = testref$Treatment_Results, positive = "Success")


## xgbTree 
# All features. (ACC, SENS, SPEC - 67%, 23%, 87%)
# Feature selec. (ACC, SENS, SPEC - 76%, 38%, 93%)
set.seed(70)
xgbTreemodel <- predict(model_list1$xgbTree, newdata = test, type="raw")
model_list1$xgbTree #75% training accuracy
confusionMatrix(data = xgbTreemodel, reference = testref$Treatment_Results, positive = "Success")

## glmboost 
# All features. (ACC, SENS, SPEC - 67%, 23%, 87%)
# Feature selec. (ACC, SENS, SPEC - 77%, 42%, 93%)
set.seed(70)
glmboostmodel <- predict(model_list1$glmboost, newdata = test, type="raw")
model_list1$glmboost #77% training accuracy
confusionMatrix(data = glmboostmodel, reference = testref$Treatment_Results, positive = "Success")


## svmLinear
# All features. (ACC, SENS, SPEC - 67%, 23%, 87%)
# Feature selec. (ACC, SENS, SPEC - 77%, 42%, 93%)
set.seed(70)
svmmodel <- predict(model_list1$svmLinear, newdata = test, type="raw")
model_list1$svmLinear #77% training accuracy
confusionMatrix(data = svmmodel, reference = testref$Treatment_Results, positive = "Success")


## ada
# All features. (ACC, SENS, SPEC - 67%, 23%, 87%)
# Feature selec. (ACC, SENS, SPEC - 76%, 28%, 97%)
set.seed(70)
adamodel <- predict(model_list1$ada, newdata = test, type="raw")
model_list1$ada #76% training accuracy
confusionMatrix(data = adamodel, reference = testref$Treatment_Results, positive = "Success")


## KNN algo
# All features. (ACC, SENS, SPEC - 73%, 19%, 97%)
# Feature selec. (ACC, SENS, SPEC - 82%, 42%, 100%)
set.seed(70)
KNNalgo9 <- train(Treatment_Results ~ protocole_1+arm_1+age+sex_1+morphine_1, data=train, method="knn", trControl = control,
                  tuneGrid=data.frame(k=seq(2,10,by=1)), metric = "Accuracy")
KNNalgo9res <- predict(KNNalgo9, newdata = test, type="raw")
KNNalgo9
plot(KNNalgo9)
confusionMatrix(data = KNNalgo9res, reference = testref$Treatment_Results, positive = "Success")
```

# Introduction 

The dataframes needed for the baseline algorithm are already created previously :

- TableAgg from Devoir 1 for patients with baseline measurements.
- TableSuc from Devoir 2 for patients that met success criteria. 

What is needed is merging them into one dataframe and preprocessing it to apply different ML models for our treatement results prediction.

\
\


# Objective 01 : Importing the dataframes and merging them.

Importing the necessary libraries

```{r, eval=F, echo=T}
require(tidyverse)
require(caret)
require(caretEnsemble)
require(missForest)
require(doParallel)
require(explore)
require(plyr)
require(janitor)
require(lfactors)
require(magrittr)
require(skimr)
require(tibble)
require(stringr)
require(kableExtra)
require(prettydoc)
require(rmdformats)
```

\
Importing the dataframes from Devoir 1 and Devoir 2

```{r Importing tables, eval=F, echo=T}
### DM Table with variable that indicates the ARM. LB Table that indicates drug results and day.
dm <- read.csv("ascii-data-files nida-ctn-0001-20230318/dm.csv", header = T, dec = ".")
dm2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/dm.csv", header = T, dec = ".")
lb <- read.csv("ascii-data-files nida-ctn-0001-20230318/lb.csv", header = T, dec = ".")
lb2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/lb.csv", header = T, dec = ".")

#Criteria of success D13 or D14 with negative opoids results.
# Importing lb database containg opioid test results, and dm with arm.
dm3 <- dm %>% filter(ARM %in% c("CLONIDINE","BUPRENORPHINE/NALOXONE")) %>% select(USUBJID, ARM)
dm4 <- dm2 %>% filter(ARM %in% c("CLONIDINE","BUPRENORPHINE/NALOXONE")) %>% select (USUBJID, ARM)
lb3 <- lb %>% filter(EPOCH == "ACTIVE", VISITNUM %in% c(1:14)) #Missing 5 patients from lb3 compared to dm3 when applying this filter. as they are still in screening epoch but have already been assigned in dm database "01_018076" "01_027452" "01_058918" "01_063935" "01_067193".
lb4 <- lb2 %>% filter(EPOCH == "ACTIVE", VISITNUM %in% c(1:14)) #Missing 27 patients from lb4 compared to dm4 when applying this filter.

# Adding the ARM column.
list_df10 = list(lb3, dm3)
tabURINE <- list_df10 %>% 
  reduce(full_join, by='USUBJID') %>% 
  select("USUBJID", "LBTEST", "LBORRES", "LBSTAT", "LBREASND", "LBSPEC", "LBMETHOD", "VISIT", "VISITNUM", "LBDY", "LBDTC", "ARM") #113 patients

list_df11 = list(lb4, dm4)
tabURINE2 <- list_df11 %>% 
  reduce(full_join, by='USUBJID') %>% 
  select("USUBJID", "LBTEST", "LBORRES", "LBSTAT", "LBREASND", "LBSPEC", "LBMETHOD", "VISIT", "VISITNUM", "LBDY", "LBDTC", "ARM") #231 patients

# Merging both tables
TableURINE <- rbind(tabURINE, tabURINE2)

# Success criteria met for these patients with NEGATIVE test results, others are a Failure.
TableSuc <- TableURINE %>% 
  filter(VISITNUM %in% c(13,14) ) %>% 
  filter(LBTEST == "OPIATE") %>% 
  filter(LBORRES == "NEGATIVE") %>% 
  mutate(PROTOCOLE = if_else(str_detect(USUBJID, "^01"), "In-patient protocol", "Out-patient protocol")) %>%
  mutate(PROTOCOLE = as.factor(PROTOCOLE)) %>% 
  filter(ARM %in% c("CLONIDINE","BUPRENORPHINE/NALOXONE")) %>% 
  # Issue with patient 02_036130. Filtered again because after join there was one patient that had SCREEN FAILURE on DM4 but got to the final table through LB4 because he had active epoch tests, which doesn't make sense.
  arrange(VISITNUM) %>% 
  select(USUBJID, LBORRES)

# TableAgg from Devoir 1, for more information please refer to Devoir 1.
dm <- read.csv("ascii-data-files nida-ctn-0001-20230318/dm.csv", header = T, dec = ".")
tabDM <- dm %>% select(USUBJID, ARM, AGE, SEX)
vs <- read.csv("ascii-data-files nida-ctn-0001-20230318/vs.csv", header = T, dec = ".")
tabTEMP <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="TEMP") %>% select(USUBJID, VSORRES)
tabTEMP <- distinct(tabTEMP)
colnames(tabTEMP) <- c("USUBJID", "TEMP")
tabRESP <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="RESP") %>% select(USUBJID, VSORRES)
tabRESP <- distinct(tabRESP)
colnames(tabRESP) <- c("USUBJID", "RESP")
tabPULSEsitting <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEsitting <- distinct(tabPULSEsitting)
colnames(tabPULSEsitting) <- c("USUBJID", "PULSEsitting")
tabPULSEstanding <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEstanding <- distinct(tabPULSEstanding)
colnames(tabPULSEstanding) <- c("USUBJID", "PULSEstanding")
tabSBPsitting <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPsitting <- distinct(tabSBPsitting)
colnames(tabSBPsitting) <- c("USUBJID", "SBPsitting")
tabSBPstanding <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPstanding <- distinct(tabSBPstanding)
colnames(tabSBPstanding) <- c("USUBJID", "SBPstanding")
tabDBPsitting <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPsitting <- distinct(tabDBPsitting)
colnames(tabDBPsitting) <- c("USUBJID", "DBPsitting")
tabDBPstanding <- vs %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPstanding <- distinct(tabDBPstanding)
colnames(tabDBPstanding) <- c("USUBJID", "DBPstanding")
tabHEIGHT <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="HEIGHT") %>% select(USUBJID, VSORRES)
tabHEIGHT <- distinct(tabHEIGHT)
colnames(tabHEIGHT) <- c("USUBJID", "HEIGHT") 
tabWEIGHT <- vs %>% filter(EPOCH=="SCREENING", VSTESTCD=="WEIGHT") %>% select(USUBJID, VSORRES)
tabWEIGHT <- distinct(tabWEIGHT)
colnames(tabWEIGHT) <- c("USUBJID", "WEIGHT")
calculate_bmi <- function(dataframe, height_col, weight_col, bmi_col) {
  height_meters <- dataframe[[height_col]] * 0.0254
  weight_kg <- dataframe[[weight_col]] * 0.453592 
  bmi <- weight_kg / (height_meters^2)
  dataframe[[bmi_col]] <- bmi 
  return(dataframe)
}
tabBMI <- left_join(tabHEIGHT, tabWEIGHT, by="USUBJID")
tabBMI <- calculate_bmi(tabBMI, "HEIGHT", "WEIGHT", "BMI") %>% select("USUBJID", "BMI")
list_df1 = list(tabBMI, tabDBPsitting, tabSBPsitting, tabDBPstanding, tabSBPstanding, tabPULSEsitting, tabPULSEstanding, tabRESP, tabTEMP)
tabVS <- list_df1 %>% reduce(full_join, by='USUBJID')
sc <- read.csv("ascii-data-files nida-ctn-0001-20230318/sc.csv", header = T, dec = ".")
tabMARITAL <- sc %>% filter(SCTESTCD=="MARITAL") %>% select(USUBJID, SCORRES)
tabMARITAL <- distinct(tabMARITAL)
colnames(tabMARITAL) <- c("USUBJID", "MARITAL")
tabEDUC <- sc %>% filter(SCTESTCD=="EDUCYRS") %>% select(USUBJID, SCORRES)
tabEDUC <- distinct(tabEDUC)
colnames(tabEDUC) <- c("USUBJID", "EDUCYRS")
tabEMPLOY30 <- sc %>% filter(SCTESTCD=="EMPLOY30") %>% select(USUBJID, SCORRES)
tabEMPLOY30 <- distinct(tabEMPLOY30)
colnames(tabEMPLOY30) <- c("USUBJID", "EMPLOY30")
tabEMPLOY3Y <- sc %>% filter(SCTESTCD=="EMPLOY3Y") %>% select(USUBJID, SCORRES)
tabEMPLOY3Y <- distinct(tabEMPLOY3Y)
colnames(tabEMPLOY3Y) <- c("USUBJID", "EMPLOY3Y")
tabEMPLOY <- left_join(tabEMPLOY30, tabEMPLOY3Y, by="USUBJID")
list_df2 = list(tabEDUC, tabEMPLOY, tabMARITAL)
tabSC <- list_df2 %>% reduce(full_join, by='USUBJID')
lb <- read.csv("ascii-data-files nida-ctn-0001-20230318/lb.csv", header = T, dec = ".")
tabMETHADONE <- lb %>% filter(EPOCH=="SCREENING", LBTESTCD=="METHADON", LBMETHOD=="CENTRAL LAB") %>% select(USUBJID, LBORRES)
tabMETHADONE <- distinct(tabMETHADONE)
colnames(tabMETHADONE) <- c("USUBJID", "METHADONE")
tabMORPHINE <- lb %>% filter(EPOCH=="SCREENING", LBTESTCD=="MORPHINE") %>% select(USUBJID, LBORRES)
tabMORPHINE <- distinct(tabMORPHINE)
colnames(tabMORPHINE) <- c("USUBJID", "MORPHINE")
list_df3 = list(tabMETHADONE, tabMORPHINE)
tabLB <- list_df3 %>% reduce(full_join, by='USUBJID')
qs <- read.csv("ascii-data-files nida-ctn-0001-20230318/qs.csv", header = T, dec = ".")
tabQS <- qs %>% filter(EPOCH=="SCREENING", QSTESTCD=="VAS001") %>% select(USUBJID, QSORRES)
tabQS <- distinct(tabQS)
colnames(tabQS) <- c("USUBJID", "VAS001")
list_df4 = list(tabDM, tabSC, tabVS, tabQS, tabLB)
tabALL_01 <- list_df4 %>% reduce(full_join, by='USUBJID')
dm2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/dm.csv", header = T, dec = ".")
tabDM2 <- dm2 %>% select(USUBJID, ARM, AGE, SEX)
vs2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/vs.csv", header = T, dec = ".")
tabTEMP2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="TEMP") %>% select(USUBJID, VSORRES)
tabTEMP2 <- distinct(tabTEMP2)
colnames(tabTEMP2) <- c("USUBJID", "TEMP")
tabTEMP2$TEMP[tabTEMP2$TEMP == 999.9] <- 99.9 
tabRESP2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="RESP") %>% select(USUBJID, VSORRES)
tabRESP2 <- distinct(tabRESP2)
colnames(tabRESP2) <- c("USUBJID", "RESP")
tabPULSEsitting2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEsitting2 <- distinct(tabPULSEsitting2)
colnames(tabPULSEsitting2) <- c("USUBJID", "PULSEsitting")
tabPULSEstanding2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="PULSE") %>% select(USUBJID, VSORRES)
tabPULSEstanding2 <- distinct(tabPULSEstanding2)
colnames(tabPULSEstanding2) <- c("USUBJID", "PULSEstanding")
tabSBPsitting2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPsitting2 <- distinct(tabSBPsitting2)
colnames(tabSBPsitting2) <- c("USUBJID", "SBPsitting")
tabSBPstanding2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="SBP") %>% select(USUBJID, VSORRES)
tabSBPstanding2 <- distinct(tabSBPstanding2)
colnames(tabSBPstanding2) <- c("USUBJID", "SBPstanding")
tabDBPsitting2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="SITTING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPsitting2 <- distinct(tabDBPsitting2)
colnames(tabDBPsitting2) <- c("USUBJID", "DBPsitting")
tabDBPstanding2 <- vs2 %>% filter(EPOCH=="SCREENING", VSPOS=="STANDING", VSTESTCD=="DBP") %>% select(USUBJID, VSORRES)
tabDBPstanding2 <- distinct(tabDBPstanding2)
colnames(tabDBPstanding2) <- c("USUBJID", "DBPstanding")
tabHEIGHT <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="HEIGHT") %>% select(USUBJID, VSORRES)
tabHEIGHT <- distinct(tabHEIGHT)
colnames(tabHEIGHT) <- c("USUBJID", "HEIGHT")
tabWEIGHT <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="WEIGHT") %>% select(USUBJID, VSORRES)
tabWEIGHT <- distinct(tabWEIGHT)
colnames(tabWEIGHT) <- c("USUBJID", "WEIGHT")
tabHEIGHT2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="HEIGHT") %>% select(USUBJID, VSORRES)
tabHEIGHT2 <- distinct(tabHEIGHT2)
colnames(tabHEIGHT2) <- c("USUBJID", "HEIGHT")
tabHEIGHT2$HEIGHT[tabHEIGHT2$HEIGHT == 715] <- 71
tabWEIGHT2 <- vs2 %>% filter(EPOCH=="SCREENING", VSTESTCD=="WEIGHT") %>% select(USUBJID, VSORRES)
tabWEIGHT2 <- distinct(tabWEIGHT2)
colnames(tabWEIGHT2) <- c("USUBJID", "WEIGHT")
calculate_bmi <- function(dataframe, height_col, weight_col, bmi_col) {
  height_meters <- dataframe[[height_col]] * 0.0254
  weight_kg <- dataframe[[weight_col]] * 0.453592 
  bmi <- weight_kg / (height_meters^2)  
  dataframe[[bmi_col]] <- bmi
  return(dataframe)
}
tabBMI2 <- left_join(tabHEIGHT2, tabWEIGHT2, by="USUBJID")
tabBMI2 <- calculate_bmi(tabBMI2, "HEIGHT", "WEIGHT", "BMI") %>% select("USUBJID", "BMI")
list_df5 = list(tabBMI2, tabDBPstanding2, tabSBPstanding2, tabDBPsitting2, tabSBPsitting2, tabPULSEstanding2, tabPULSEsitting2, tabRESP2, tabTEMP2)
tabVS2 <- list_df5 %>% reduce(full_join, by='USUBJID')
sc2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/sc.csv", header = T, dec = ".")
tabMARITAL2 <- sc2 %>% filter(SCTESTCD=="MARITAL") %>% select(USUBJID, SCORRES)
tabMARITAL2 <- distinct(tabMARITAL2)
colnames(tabMARITAL2) <- c("USUBJID", "MARITAL")
tabEDUC2 <- sc2 %>% filter(SCTESTCD=="EDUCYRS") %>% select(USUBJID, SCORRES)
tabEDUC2 <- distinct(tabEDUC2)
colnames(tabEDUC2) <- c("USUBJID", "EDUCYRS")
tabEMPLOY302 <- sc2 %>% filter(SCTESTCD=="EMPLOY30") %>% select(USUBJID, SCORRES)
tabEMPLOY302 <- distinct(tabEMPLOY302)
colnames(tabEMPLOY302) <- c("USUBJID", "EMPLOY30")
tabEMPLOY3Y2 <- sc2 %>% filter(SCTESTCD=="EMPLOY3Y") %>% select(USUBJID, SCORRES)
tabEMPLOY3Y2 <- distinct(tabEMPLOY3Y2)
colnames(tabEMPLOY3Y2) <- c("USUBJID", "EMPLOY3Y")
tabEMPLOY2 <- left_join(tabEMPLOY302, tabEMPLOY3Y2, by="USUBJID")
list_df6 = list(tabEDUC2, tabEMPLOY2, tabMARITAL2)
tabSC2 <- list_df6 %>% reduce(full_join, by='USUBJID')
lb2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/lb.csv", header = T, dec = ".")
tabMETHADONE2 <- lb2 %>% filter(EPOCH=="SCREENING", LBTESTCD=="METHADON", LBMETHOD=="CENTRAL LAB") %>% select(USUBJID, LBORRES)
tabMETHADONE2 <- distinct(tabMETHADONE2)
colnames(tabMETHADONE2) <- c("USUBJID", "METHADONE")
tabMORPHINE2 <- lb2 %>% filter(EPOCH=="SCREENING", LBTESTCD=="MORPHINE") %>% select(USUBJID, LBORRES)
tabMORPHINE2 <- distinct(tabMORPHINE2)
colnames(tabMORPHINE2) <- c("USUBJID", "MORPHINE") 
list_df7 = list(tabMETHADONE2, tabMORPHINE2)
tabLB2 <- list_df7 %>% reduce(full_join, by='USUBJID')
qs2 <- read.csv("ascii-data-files nida-ctn-0002-20230318/qs.csv", header = T, dec = ".")
tabQS2 <- qs2 %>% filter(EPOCH=="SCREENING", QSTESTCD=="VAS001") %>% select(USUBJID, QSORRES)
tabQS2 <- distinct(tabQS2)
colnames(tabQS2) <- c("USUBJID", "VAS001")
list_df8 = list(tabDM2, tabSC2, tabVS2, tabQS2, tabLB2)
tabALL_02 <- list_df8 %>% reduce(full_join, by='USUBJID')
TableALL <- rbind(tabALL_02, tabALL_01)
TableALL$VAS001 <- as.numeric(TableALL$VAS001)

TableAgg <- TableALL %>%
  group_by(USUBJID, ARM, SEX, EDUCYRS, EMPLOY30, EMPLOY3Y, MARITAL, METHADONE, MORPHINE) %>%
  dplyr::summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE))) %>% 
  mutate(dplyr::across(everything(), ~replace_na(.x, NA)))

TableAgg <- distinct(TableAgg) %>% 
  filter(!(ARM == "")) %>% 
  filter(!(ARM == "SCREEN FAILURE"))
TableAgg$ARM <- as.factor(TableAgg$ARM)
TableAgg$SEX <- as.factor(TableAgg$SEX)
TableAgg$EDUCYRS <- as.numeric(TableAgg$EDUCYRS)
TableAgg$EMPLOY30 <- as.factor(TableAgg$EMPLOY30)
TableAgg$EMPLOY3Y <- as.factor(TableAgg$EMPLOY3Y)
TableAgg$MARITAL <- as.factor(TableAgg$MARITAL)
TableAgg$METHADONE <- as.factor(TableAgg$METHADONE)
TableAgg$MORPHINE <- as.factor(TableAgg$MORPHINE)
TableAgg <- as.data.frame(TableAgg) 
```

\
Merging the dataframes.

```{r Merging tables, eval=F, echo=T}
# Merging this table with the initial table we have from Devoir 1 
# Now we have the table with Success 1 or Failure 0 in LBORRES column.
list_df20 = list(TableAgg, TableSuc)
Dataset_01 <- list_df20 %>% 
  reduce(full_join, by='USUBJID') %>% 
  mutate(PROTOCOLE = if_else(str_detect(USUBJID, "^01"), "In-patient protocol", "Out-patient protocol")) %>%
  mutate(PROTOCOLE = as.factor(PROTOCOLE)) %>% 
  mutate(LBORRES = if_else(is.na(LBORRES), "0", LBORRES)) %>% 
  mutate(LBORRES = if_else(str_detect(LBORRES, "NEGATIVE"), "1", "0")) %>%
  mutate(LBORRES = as.factor(LBORRES)) %>% 
  mutate(EDUCYRS = as.factor(EDUCYRS))
```

```{r Merged df, echo=FALSE}
head(Dataset_01)
```

\
\

# Objective 02 : Exploratory data analysis (EDA).

```{r EDA, echo=TRUE}
#explore(Dataset_01)
#Very useful function that doesn't work unfortunately with Rmarkdown.
```

\
\

# Objective 03 : Preprocessing the dataframe.

```{r Preprocessing, eval=F, echo=T}
# Preparing the dataset to apply ML algorithms.
# filling NA, One hot encoding, Normalizing
skim(Dataset_01)
Dataset_01_treated <- Dataset_01 %>% 
  select(LBORRES, USUBJID, PROTOCOLE, ARM, AGE, SEX, BMI, METHADONE, MORPHINE, VAS001, EDUCYRS, EMPLOY30, EMPLOY3Y, MARITAL, DBPsitting, SBPsitting, PULSEsitting, RESP, TEMP) %>% 
  mutate(PROTOCOLE = if_else(str_detect(PROTOCOLE, "^In"), "0", "1")) %>% 
  mutate(PROTOCOLE = as.factor(PROTOCOLE)) %>% 
  mutate(ARM = if_else(str_detect(ARM, "^BUP"), "0", "1")) %>% 
  mutate(ARM = as.factor(ARM)) %>% 
  mutate(SEX = if_else(str_detect(SEX, "F"), "0", "1")) %>% 
  mutate(SEX = as.factor(SEX)) %>% 
  mutate(METHADONE = if_else(str_detect(METHADONE, "^NEG"), "0", "1")) %>%
  mutate(METHADONE = as.factor(METHADONE)) %>% 
  mutate(MORPHINE = if_else(str_detect(MORPHINE, "^NEG"), "0", "1")) %>% 
  mutate(MORPHINE = as.factor(MORPHINE))

# Removing NA # https://www.kaggle.com/code/lmorgan95/missforest-the-best-imputation-algorithm/report
set.seed(70)
Dataset_IMP <- subset(Dataset_01_treated, select = -c(LBORRES, USUBJID))
imp_missForest <- missForest(Dataset_IMP, maxiter = 10, verbose = T)
Dataset_IMP <- imp_missForest$ximp

# USUBJID Not needed as we already extracted arm information in a seperate column
LBORRES <- Dataset_01_treated[, 1]
Dataset_01_treated <- cbind(LBORRES, Dataset_IMP)

# Normalizing values 
preproc <- preProcess(Dataset_01_treated, method = "range")
DS01_norm <- predict(preproc, Dataset_01_treated)

# Dummy variables
dmy <- dummyVars(" ~ .", data = DS01_norm, fullRank=T)
DS01_norm_dummy <- data.frame(predict(dmy, newdata = DS01_norm)) #FullRank to choose one dummy for columns that has 2 values, ex GENDER.
DS01_norm_dummy <- DS01_norm_dummy %>% clean_names()
colnames(DS01_norm_dummy)[colnames(DS01_norm_dummy) == "lborres_1"] <- "Treatment_Results"
DS01_norm_dummy$Treatment_Results <- lfactor(DS01_norm_dummy$Treatment_Results, levels = 0:1, labels = c("Failure", "Success"))
DS01_norm_dummy$Treatment_Results <- as.factor(DS01_norm_dummy$Treatment_Results)

set.seed(70)
index <- createDataPartition(DS01_norm_dummy$Treatment_Results, p = 0.80, list = FALSE)
train <- DS01_norm_dummy[index,]
test <- DS01_norm_dummy[-index,]
testref <- DS01_norm_dummy[-index,]

TableSKIMR <- DS01_norm_dummy %>% 
  skim() %>% 
  dplyr::select(-skim_type)
```

```{r Ready dataset for ML, echo=TRUE}
str(DS01_norm_dummy)
```

\
\

# Objective 04 : Applying ML models

## Baseline accuracy

If our prediction are all "Failure" we get an Accuracy of 69% (Sens 0 no positive class). This should be the baseline we improve upon in both Accuracy and Sensitivity.

```{r, echo=TRUE}
test$Treatment_Results = "Failure"
test$Treatment_Results <- lfactor(test$Treatment_Results, levels = 0:1, labels = c("Failure", "Success"))
confusionMatrix(data = test$Treatment_Results, reference = testref$Treatment_Results, positive = "Success")
```

\

## Feature selection

### RFE

```{r echo=TRUE, warning=FALSE}
## Feature selection using RFE (Recursive Feature Elimination)
# Selected variable : ARM, PROTOCOLE
set.seed(70)
Control <- rfeControl(functions=rfFuncs, method="cv", number=10) 
Results <- rfe(DS01_norm[,2:18], DS01_norm[,1], sizes=c(1:16), metric = "RMSE", rfeControl=Control)
print(Results)
```

```{r echo=FALSE}
predictors(Results)
```

### SBF

```{r, echo=TRUE}
## Feature selection using SBF (Selection By Filtering)
# Selected variable : ARM (100%), MORPHINE (100%), PROTOCOLE (100%), SBPsitting (100%), PULSEsitting (60%)
RFwithGAM <- sbf(numeric_cols[,2:14], numeric_cols[,1],
                 sbfControl = sbfControl(functions = rfSBF,
                                         verbose = FALSE,
                                         method = "cv"))
print(RFwithGAM)
```

\
\

## Model selection
  
```{r, echo=TRUE}
# Training multiple models
number <- 5
repeats <- 4
control <- trainControl(method = "cv", 
                        classProbs = TRUE, 
                        savePredictions = "final",
                        index = createResample(train$Treatment_Results, repeats*number), 
                        summaryFunction = multiClassSummary, 
                        allowParallel = TRUE)
set.seed(70)

model_list1 <- caretList(Treatment_Results ~ protocole_1+arm_1+morphine_1+sb_psitting+puls_esitting, 
                         data = train,
                         trControl = control,
                         metric = "Accuracy", 
                         methodList = my_models)

```

```{r echo=FALSE}
# Plotting with accuracy the different models: 
list_of_results <- lapply(my_models, function(x) {model_list1[[x]]$resample})
df_results <- do.call("bind_rows", list_of_results)
df_results %<>% mutate(Model = lapply(my_models, function(x) {rep(x, number*repeats)}) %>% unlist())
df_results %>% 
  select(Accuracy, Model) %>% 
  ggplot(aes(Model, Accuracy, fill = Model, color = Model)) + 
  geom_boxplot(show.legend = FALSE, alpha = 0.3) + 
  theme_minimal() + 
  coord_flip()
```
